Community-Conditioned Language Models

### 2020-10-30 paper/paper.tex
some tables
add really big table
Losses of CCLM

\documentclass[11pt,a4paper]{article}
\usepackage[hyperref]{acl2020}
\usepackage[T1]{fontenc}
\usepackage{times}
\usepackage{latexsym}
\usepackage{siunitx}
\renewcommand{\UrlFont}{\ttfamily\small}

@@

At the the core of the community-conditioned language model (CCLM) 
is a standard $n$-layer neural sequence encoder 
with token embedding (we experiment with LSTM and transformer encoders).
To condition the model, we additionally train a \emph{community embedding},
which is concatenated with the hidden layer of the sequence encoder
for some layer $l_c \leq n$ and passed through a linear layer, 
which projects the resulting vector back down to the origial hidden size.
For $l_c=0$, the community embedding is concatenated directly to the token embedding.
For this reason, the hidden size of the sequence encoder 
and the size of the token embedding are always the same.

@@

All experiments used models with \num{3} sequence encoder layers, 
each with hidden (and token embedding) size of \num{256}. 
The transformer models had \num{8} attention heads per layer.\footnote{
  This number of attention heads was chosen to give the LSTM and transformer
  models a comparable number of paramters 
  (\num{22171203} and \num{21779523}, respectively).}
The conditional models were given a community embedding with \num{16} dimensions.

@@

@@

% Computation of loss/community/model (show all)
\begin{table}
  \centering
  \begin{tabular}{ccc}
    todo &todo &todo 
  \end{tabular}
  \caption{Model losses per community}
  \label{tab:losses}
\end{table}
TODO commentary. Why are some communities low/high.

>>>>

\documentclass[11pt,a4paper]{article}
\usepackage[hyperref]{acl2020}
\usepackage[T1]{fontenc}
\usepackage{times}
\usepackage{latexsym}
\usepackage{siunitx}
\usepackage{booktabs}
\renewcommand{\UrlFont}{\ttfamily\small}

@@

At the the core of the community-conditioned language model (CCLM) 
is a standard $n$-layer neural sequence encoder 
with token embedding (we experiment with LSTM and transformer encoders).
To condition the model, we additionally train a \emph{community embedding},
which is concatenated with the hidden layer of the sequence encoder
for some layer $c \leq n$ and passed through a linear layer, 
which projects the resulting vector back down to the origial hidden size.
For $c = n$, the output of this linear layer is passed directly to the softmax function,
just as the final hidden layer of the sequence encoder is in other models.
For $c=0$, the community embedding is concatenated with the token embedding.
For this reason, we set the hidden size of the sequence encoder 
and the size of the token to be embedding equal for all models.

@@

All experiments used models with \num{3} sequence encoder layers, 
each with hidden (and token embedding) size of \num{256}. 
The transformer models had \num{8} attention heads per layer.\footnote{
  This number of attention heads was chosen in part to give the LSTM and transformer
  models a comparable number of paramters 
  (\num{22171203} and \num{21779523}, respectively).}
The conditional models were given a community embedding with \num{16} dimensions.

@@

\begin{table}
  \centering
  \input{floats/best_epoch.tex}
  \label{tab:best-epoch}
  \caption{Epoch with the lowest validation loss.}
\end{table}

@@

As one would expect, 
the conditioned models have lower perplexity 
than their respective unconditioned baseline models,
but this effect is different in magnitude
for models with different community embedding depths .
For the LSTM encoder, 
the best model ($c=1$) concatenates the community embedding after the first encoder layer,
with two more LSTM layers after that.
The best transformer model ($c=3$), on the other hand,
incorporates the community information last,
after all the sequence encoder layers
(Table \ref{tab:model-ppls}).

TODO: For space reasons, it might be better to put the community-level table in the appendix
and have the table in body just show the means (and perhaps medians/stddevs?) for each architecture.

We also note that the conditioned models perform differently across different communities---\
for some communities, the test perplexity of the conditioned models is a lot lowe, 
whereas for others it is the same or even slightly higher.
(TODO: talk about this in terms of mutual information instead?)
As an anecdotal observation, we note 
that communities where conditioning has the least effect
tend to be more organized around more ``general interest'' topics.
This makes sense intuitively, since communities with more niche topics
would tend to have more specialized vocabulary and speech patterns.

The pattern of how much the perplexity falls across different communities
is similar across architectures, but with some differences.
It could be interesting to further investigate these differences in future work,
since it could reveal differences in the kind of linguistic variation the different
model architectures capture.

\begin{table*}
  \footnotesize
  \centering
  \input{floats/model_ppls.tex}
  \caption{Mean model test perplexity per community.}
  \label{tab:model-ppls}
\end{table*}
